---
title: "p8105_hw6_tp2859"
author: "Tingcheng Pan"
date: "2024-12-02"
output: html_document
---
#Problem1
```{r}
# Load required libraries
library(dplyr)
library(ggplot2)
library(modelr)
library(broom)
library(purrr)

set.seed(123)
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
# Step 1: Fit the initial regression model
initial_model <- lm(tmax ~ tmin, data = weather_df)

# Step 2: Bootstrap sampling (5000 samples)
bootstrap_results <- weather_df %>%
  modelr::bootstrap(n = 5000) %>%
  mutate(
    model = map(strap, ~ lm(tmax ~ tmin, data = .x)),
    glance = map(model, broom::glance),
    tidy = map(model, broom::tidy),
    r_squared = map_dbl(glance, "r.squared"),
    beta0_beta1_log = map_dbl(tidy, ~ {
      beta0 <- .x %>% filter(term == "(Intercept)") %>% pull(estimate)
      beta1 <- .x %>% filter(term == "tmin") %>% pull(estimate)
      if (!is.na(beta0) && !is.na(beta1)) {
        return(log(beta0 * beta1))
      } else {
        return(NA)  # Handle cases where beta0 or beta1 is missing
      }
    })
  )

# Step 3: Handle NaN and missing values
# Check for NA or NaN values
num_na_r_squared <- sum(is.na(bootstrap_results$r_squared))
num_na_beta0_beta1_log <- sum(is.na(bootstrap_results$beta0_beta1_log))

print(paste("Number of NA in r_squared:", num_na_r_squared))
print(paste("Number of NA in beta0_beta1_log:", num_na_beta0_beta1_log))

# Exclude NaN values when calculating quantiles
r_squared_ci <- quantile(bootstrap_results$r_squared, c(0.025, 0.975), na.rm = TRUE)
beta0_beta1_log_ci <- quantile(bootstrap_results$beta0_beta1_log, c(0.025, 0.975), na.rm = TRUE)

# Print confidence intervals
print("95% CI for r^2:")
print(r_squared_ci)
print("95% CI for log(beta0 * beta1):")
print(beta0_beta1_log_ci)

# Step 4: Visualize the bootstrap distributions
bootstrap_estimates <- bootstrap_results %>%
  select(r_squared, beta0_beta1_log)

# Plot for r_squared
ggplot(bootstrap_estimates, aes(x = r_squared)) +
  geom_histogram(bins = 30, fill = "blue", alpha = 0.6, color = "black") +
  labs(
    title = "Bootstrap Distribution of r^2",
    x = "r^2",
    y = "Frequency"
  ) +
  theme_minimal()

# Plot for log(beta0 * beta1)
ggplot(bootstrap_estimates, aes(x = beta0_beta1_log)) +
  geom_histogram(bins = 30, fill = "green", alpha = 0.6, color = "black") +
  labs(
    title = "Bootstrap Distribution of log(beta0 * beta1)",
    x = "log(beta0 * beta1)",
    y = "Frequency"
  ) +
  theme_minimal()
```


#Problem2
```{r}
# Load required libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(broom)
library(purrr)
library(tidyverse)
library(dplyr)
library(mgcv)
library(modelr)
library(SemiPar)

# Load the dataset
homicide_data <- read.csv("homicide-data.csv")
# Step 1: Create city_state variable
homicide_data <- homicide_data %>%
  mutate(
    city_state = paste(city, state, sep = ", "),
    resolved = ifelse(disposition %in% c("Closed by arrest"), 1, 0),
    victim_age = as.numeric(victim_age)
  )
```
```{r}
# Step 2: Clean the data
homicide_data_clean <- homicide_data %>%
  # Remove specified cities
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")
  ) %>%
  # Restrict to victim races "White" or "Black"
  filter(
    victim_race %in% c("White", "Black")
  ) %>%
  # Remove rows with missing victim_age and ensure it's numeric
  filter(!is.na(victim_age),
         !is.na(lat),
         !is.na(lon)) %>%
  mutate(
    victim_age = as.numeric(victim_age)
  )
```

```{r}
# Step 3: Logistic regression for Baltimore, MD
baltimore_data <- filter(homicide_data_clean, city_state == "Baltimore, MD")

# Fit logistic regression model
baltimore_model <- glm(
  resolved ~ victim_age + victim_sex + victim_race,
  data = baltimore_data,
  family = "binomial"
)

# Extract adjusted odds ratios and confidence intervals
baltimore_results <-broom::tidy(baltimore_model, exponentiate = TRUE, conf.int = TRUE)

baltimore_results %>%
  filter(term == "victim_sexMale") %>%
  select(estimate, conf.low, conf.high)
```

```{r}
# Step 4: Logistic regression for all cities
city_models <- homicide_data_clean %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~ glm(
      resolved ~ victim_age + victim_sex + victim_race,
      family = binomial,
      data = .x
    )),
    results = map(model, ~ broom::tidy(
      .x,
      exponentiate = TRUE,
      conf.int = TRUE
    ))
  ) %>%
  unnest(results) %>%
  filter(term == "victim_sexMale") # Focus on male vs female comparison
city_models
```
```{r}
# Step 5: Create a plot of the Odds Ratios
ggplot(city_models, aes(x = reorder(city_state, estimate), y = estimate)) +
  geom_point(color = "blue") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.4, color = "blue") +
  coord_flip() +
  labs(
    title = "Odds Ratios for Solving Homicides (Male vs Female Victims)",
    x = "City",
    y = "Odds Ratio (Male vs Female Victims)"
  ) +
  theme_minimal()

```
The plot provides an insightful visual representation of gender-related differences in homicide resolution across cities. While most cities show no significant disparity (OR ≈ 1), a few cities—Albuquerque, NM; Stockton, CA; Fresno, CA; New York, NY; Baton Rouge, LA; and Cincinnati, OH—exhibit notable deviations, warranting further investigation into local factors affecting homicide resolution.

#Problem3
```{r}
# Load the dataset
birthweight <- read.csv("birthweight.csv")

# Step 1: Data Cleaning and Preparation
# Update data cleaning with the correct number of labels
birthweight <- birthweight %>%
  mutate(
    babysex = factor(babysex, labels = c("Male", "Female")),
    malform = factor(malform, labels = c("Absent", "Present")),
    frace = case_when(
      frace == 1 ~ "White",
      frace == 2 ~ "Black",
      frace == 3 ~ "Asian",
      frace == 4 ~ "Puerto Rican",
      frace == 8 ~ "Other",
      frace == 9 ~ "Unknown",
      TRUE ~ "Invalid" # For unexpected values
    ) %>% factor(),
    mrace = case_when(
      mrace == 1 ~ "White",
      mrace == 2 ~ "Black",
      mrace == 3 ~ "Asian",
      mrace == 4 ~ "Puerto Rican",
      mrace == 8 ~ "Other",
      mrace == 9 ~ "Unknown",
      TRUE ~ "Invalid" # For unexpected values
    ) %>% factor(),
    parity = as.factor(parity)
  )%>%
  drop_na()
```
The model I proposed is using following data:
Gestational Age (gaweeks): Longer gestational age is strongly associated with higher birthweight.
Mother's Weight Gain (wtgain): Increased weight gain during pregnancy is linked to better fetal growth.
Mother's Pre-pregnancy BMI (ppbmi): Maternal nutritional status can significantly impact fetal growth.
Mother's Race (mrace): Maternal race/ethnicity often reflects socioeconomic and health disparities affecting birthweight.
Presence of Malformations (malform): Structural malformations may reduce birthweight.
Mother's Height (mheight): Taller mothers tend to have larger infants due to their larger body frames.
```{r}
# Step 2: Propose a Primary Model
# Include predictors: mother's weight gain, gestational age, and mother's height
set.seed(123)
proposed_model <- lm(
  bwt ~ gaweeks + wtgain + ppbmi + mrace + malform + mheight,
  data = birthweight
)
model1 <- lm(bwt ~ blength + gaweeks, data = birthweight)

# Model 2: Head circumference, length, sex, and all interactions
model2 <- lm(bwt ~ bhead * blength * babysex, data = birthweight)

```

```{r}
# Step 3: Evaluate Models with Residual Plots
birthweight <- birthweight %>%
  add_predictions(proposed_model, var = "pred_proposed") %>%
  add_residuals(proposed_model, var = "resid_proposed") %>%
  add_predictions(model1, var = "pred_model1") %>%
  add_residuals(model1, var = "resid_model1") %>%
  add_predictions(model2, var = "pred_model2") %>%
  add_residuals(model2, var = "resid_model2")

ggplot(birthweight, aes(x = pred_proposed, y = resid_proposed)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "Residuals vs. Fitted Values (Proposed Model)",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()
```

```{r}
set.seed(123)
cv_results <- crossv_mc(birthweight, 100) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble),
    proposed_model = map(train, ~ lm(bwt ~ gaweeks + wtgain + ppbmi + mrace + malform + mheight, data = .)),
    model1 = map(train, ~ lm(bwt ~ blength + gaweeks, data = .)),
    model2 = map(train, ~ lm(bwt ~ bhead * blength * babysex, data = .)),
    rmse_proposed = map2_dbl(proposed_model, test, ~ rmse(.x, .y)),
    rmse_model1 = map2_dbl(model1, test, ~ rmse(.x, .y)),
    rmse_model2 = map2_dbl(model2, test, ~ rmse(.x, .y))
  )
```

```{r}
# Step 6: Visualize RMSE Comparison
cv_long <- cv_results %>%
  select(rmse_proposed, rmse_model1, rmse_model2) %>%
  pivot_longer(cols = everything(), names_to = "Model", values_to = "RMSE")

ggplot(cv_long, aes(x = Model, y = RMSE, fill = Model)) +
  geom_boxplot() +
  labs(
    title = "Cross-Validated RMSE Comparison",
    x = "Model",
    y = "RMSE"
  ) +
  theme_minimal()
```

x

